---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# IOPAC

This is a repository containing the Input-Output model for Pacific Coast 
Fisheries (IOPAC), created by Jerry Leonard. This model estimates gross changes
in economic contributions and impacts due to changes in fishery harvests, for
example from environmental or policy changes. This readme documents the
repository and provides a minimal reproducible example.

## Running the model

Running the model can be done through the high-level wrapper function 
`iopac_wrap`. Installing the package requires `devtools`.

```{r, eval=FALSE}
library(devtools)
library(dplyr)
library(here)
here()
#The working directory should be the top level of the package.
#You can setwd("..") to install it if you want but then it might be a good idea
#to reload the project
#install("IOPAC")

#I'm trying something a little funky here so we can keep the data separate
#(allowing us to push things to a public Git).
datadir <- file.path("..", "IOPAC_data")
file_list <- list.files(datadir)
for (i in 1:length(file_list)){
  load(paste0(file.path("..", "IOPAC_data"), "/",  file_list[i]))
}

#I forgot to create a column when I was repulling the data, please excuse the
#ugly code
costflist_2017$processor$Value <- costflist_2017$processor$Xn2017
costflist_2018$processor$Value <- costflist_2018$processor$Xn2018
costflist_2019$processor$Value <- costflist_2019$processor$Xn2019
costflist_2020$processor$Value <- costflist_2020$processor$Xn2020
costflist_2021$processor$Value <- costflist_2021$processor$Xn2021
costflist_2022$processor$Value <- costflist_2022$processor$Xn2022

costflist_2017$processor$ShareC <- costflist_2017$processor$Xn2017/
  costflist_2017$processor$Xn2017[costflist_2017$processor$Type=="Revenue"]
costflist_2018$processor$ShareC <- costflist_2018$processor$Xn2018/
  costflist_2018$processor$Xn2018[costflist_2018$processor$Type=="Revenue"]
costflist_2019$processor$ShareC <- costflist_2019$processor$Xn2019/
  costflist_2019$processor$Xn2019[costflist_2019$processor$Type=="Revenue"]
costflist_2020$processor$ShareC <- costflist_2020$processor$Xn2020/
  costflist_2020$processor$Xn2020[costflist_2020$processor$Type=="Revenue"]
costflist_2021$processor$ShareC <- costflist_2021$processor$Xn2021/
  costflist_2021$processor$Xn2021[costflist_2021$processor$Type=="Revenue"]
costflist_2022$processor$ShareC <- costflist_2022$processor$Xn2022/
  costflist_2022$processor$Xn2022[costflist_2022$processor$Type=="Revenue"]
```

The wrapper function calls on eleven total inputs, two of which are optional. 
The output is a (n*m) by 12 data frame, where the first three columns 
correspond to the geographic region, sector, and name of sector for a 
multiplier, where there are _n_ regions and _m_ sectors. Then, the remaining 9 
columns correspond to the output, income, and employment multipliers for 
vessels, processors, and their aggregate (total), for each region-sector. A 
missing value denotes that there is an absence of data (e.g. fish tickets) for 
that region-sector.

```{r, eval=TRUE}
library(IOPAC)
#multres <- iopac_wrap()
#head(multres)
```

The wrapper function creates multipliers for the vessel and processor through 
the functions `make_v_mults` and `make_p_mults` respectively. These can be used 
to create individual multipliers instead if desired. Documentation for all 
functions can be found by typing ?function, for example:

```{r, eval=FALSE}
?iopac_wrap
```

Therefore, `make_v_mults` and `make_p_mults` create the multipliers, while
`iopac_wrap` passes the relevant data to the functions.

The package also includes a number of data sets that can produce multipliers 
for the West Coast region, and documentation for all included data can also be 
found similarily (typing ?data). The default data used are the most most recent, 
but the years 2018, 2019, and 2020 are included for cost, markup, and fish 
ticket data where available. 

```{r, eval=FALSE}
#?ticslist_2020
```

Therefore the user can create multipliers calibrating the model with costs or
catches from different years, or use their own data if desired, as long as the 
data matches the same format. For example, calibrating the model with cost, fish
ticket, and processor markup data from 2019 below.

```{r, eval=TRUE}

#multres <- iopac_wrap(costfin = costflist_2020, ticsin = ticslist_2020, 
#	markupsin = markups_2020)

#subset(multres, Region == 'WC' & Name == 'Whiting, Trawl')$Processor_income

```

# Adding uncertainty
I can think of at least three areas where we have a distribution for the data
but only rely on means - fish tickets, costs, and processor markups. As an
example the costs are currently the proportions each vessel spends on a
category:

```{r, eval=TRUE}
#head(costflist_2020[[1]])
```

These are the averages *n* vessels. We could pull from a normally distributed
random variable with these means and an associated standard devation? Repeat for
some number of iterations and average the results.

Fish ticket data is treated similarly although I think the processing into
proportions is done in make_v_mults currently. The same intuition would apply
although I'm unsure of the interaction between the two random variables.

```{r, eval=TRUE}
#head(ticslist_2020[[1]])
```

For both of these we could pull new data in iopac_wrap, and pass the pulled
data into make_v_mults (instead of the current static data).